---
title: "Distribution And Order Statistics"
author: "Ashik E Elahi"
date: "2024-02-17"
output: html_document
---

##### Problem: 1 - Suppose X follows a normal distribution with mean $\mu$ and variance $\sigma^2$.
##### (i) Plot the distributions on a graph paper separately with ($\mu = 3, \sigma^2 = 16), (\mu = 5, \sigma^2 = 16)$, and ($\mu = 4, \sigma^2 = 16$) and compare the graphs.

##### (ii) Plot the distributions on same graph paper with ($\mu = 4, \sigma^2 = 4), (\mu = 4, \sigma^2 = 9)$, and ($\mu = 4, \sigma^2 = 16)$


##### Solution 1 (i):

```{r}
set.seed(21)
par(mfrow = c(1, 3))
mu <- c(3, 5, 7)
sigma.2 <- c(16, 16, 16)
sigma <- sqrt(sigma.2)
colors <- c("red", "green", "blue")
labels <- paste0("N(", mu, ",", sigma, ")")
x <- seq(-20, 20, 0.01)

for(i in 1:length(mu)) {
  fx <- dnorm(x, mean = mu[i], sd = sigma[i])
  plot(x, fx, type="l", col = colors[i], lwd=3,
      main = labels[i], xlab = "x ->", 
      ylab = "Frequency f(x)")
  
  segments(x0 = mu[i], y0 = 0, x1 = mu[i], y1 = max(fx), lwd = 2, lty=3)
}
```

##### Solution 1 (ii):

```{r}
set.seed(2112)
mu <- c(4, 4, 4)
sigma.2 <- c(4, 9, 16)
sigma <- sqrt(sigma.2)
colors <- c("red", "green", "blue")
labels <- paste0("N(", mu, ",", sigma, ")")
x <- seq(-20, 20, 0.01)

for(i in 1:length(mu)) {
  fx <- dnorm(x, mean = mu[i], sd = sigma[i])
  if(i == 1) {
    plot(x, fx, type="l", col = colors[i], lwd=3,
        main = labels[i], xlab = "x ->", 
        ylab = "Frequency f(x)")
  } else {
    lines(x, fx, lwd=3, col=colors[i])
  }
}
abline(v=mu[1], lty=3) 

```

##### Problem 2: Generate 1200 random samples of size n = 12 with $\mu = 4$ and $\sigma^2 = 9$. Suppose $s^2=\frac{1}{n-1}\sum_i^n (x_i - \bar{x})^2$. Using sitable graph, justify that $\frac{(n-1)s^2}{\sigma^2}$ follows a chi-square distribution with (n-1) degress of freedom.

##### **Method - 1:**

```{r}
set.seed(221)
par(mfrow=c(1, 1))
n = 12
mu = 4
sigma = sqrt(9)
it = 1200

xvar = array()
for(i in 1:it) {
  x = rnorm(n, mu, sigma)
  xvar[i] = var(x)  
}
# xvar -> list of variances

x = ((n-1)*xvar)/sigma^2
hist(x, breaks = 15, freq = F, col = "gray", xlab = expression(((n-1)*s^2)/sigma^2), main="Chi-square distribution")

x.2 = seq(0,120, 0.01)
chisq.pdf = dchisq(x.2, df=n-1)
lines(x.2, chisq.pdf, col="blue", lwd=3)
```

##### **Method - 2:**
```{r}
set.seed(11)
#Generate 1200 random sample of size 12
n = 12
mu = 2
sigma = sqrt(9)

samples <- replicate(1200, rnorm(n, mean = mu, sd = sigma))
# samples

# apply(samples, 2, var)

#Calculate the test statistic for each sample
test_statistic <- (n-1)*(apply(samples, 2, var)/sigma^2)


hist(test_statistic, breaks = 25, probability= TRUE,
     main = "Distribution of Test Statistic",
     xlab="Test Statistic",
     ylab="Density")

# Plot the chi-square distribution with (n-1) degrees of freedom 
x <- seq(0, max(test_statistic)+5, length.out = 100)
lines(x, dchisq(x, df=n-1), col="red", lwd=2, lty=6)

#dchisq(x, df=n-1)
```

##### Problem 3: A random sample of size 4 is drawn from a population that has a uniform distribution on the interval (0, 5). The resulting order statistics are $X_{1:4}, X_{2:4}, X_{3:4}$ and $X_{4:4}$. Determine the cumulative distribution function (CDF) of the $3^{rd}$ order statistics $(X_{3:4})$ and plot the CDF of the $3^{rd}$ order statistic. Evaluate the probability $P[X_{3:4} >= 2]$


###### Method - 1:

```{r}
cdf_uniform <- function(x, a, b) {
  ifelse(x < 0, 0, ifelse(x > 5, 1, (x-a)/(b-a)))
}

Fr_n_eqn <- function(x, r, n) {
  sum_eqn <- ""
  for(j in r:n) {
    term <- paste0("(", n, "C", j, ")", "F(x)^", j, "[1-F(x)]^(",n, "-", j, ")")
    if(j < n) {
      sum_eqn = paste0(sum_eqn, term, "+")
    } else {
      sum_eqn = paste0(sum_eqn, term)
    }
  }
  return(sum_eqn)
}

Fr_n_value <- function(x, r, n, a, b) {
  cdf_x = cdf_uniform(x, a, b)
  sum_value = 0
  
  for(j in r:n) {
    sum_value = sum_value + choose(n, j)*(cdf_x^j)*(1-cdf_x)^(n-j)
  }
  return(sum_value)
}


# Results

r <- 3
n <- 4
a <- 0
b <- 5
x <- seq(a, b, length.out=100)

Fr_n_equation <- Fr_n_eqn(x, r, n)
Fr_n_values <- Fr_n_value(x, r, n, a, b)



cat("The distribution function F", r, ":", n, "(x) is given by :\n", Fr_n_equation)

plot(x, Fr_n_values, type="l", xlab = "x->", ylab = paste0("F", r, ":", n, "(x)"))

# P(x_3:4 > 2)
x_threshold <- 2
p_x3_gt_x <- 1 - Fr_n_value(x_threshold, r, n, a, b)
p_x3_gt_x
```

##### Problem 4: Suppose $X_{1:7} < X_{2:7} < X_{3:7} < X_{4:7} < X_{5:7} < X_{6:7} < X_{7:7}$ denote the order statistics based on a random sample of size 7 from the distribution having probability density function (pdf) $f(x) = \frac{1}{10}, 0 \leq x \leq 10$. Evaluate the mean and variance of the $5^{th}$ order statistic $(X_{5:7})$ and plot the CDF of the $5^{th}$ order statistic.


```{r}
cdf_dis = function(x, a, b) {
  ifelse(x < a, 0, ifelse(x > b, 1, x / 10))
}

# pdf of order statistic
fr_n = function(x, r, n, a, b) {
  cdf_x = cdf_dis(x, a, b)
  pdf_x = 0.1
  
  numerator <- factorial(n) / (factorial(r-1)*factorial(n-r))
    prob_dist <- numerator * cdf_x^(r-1) * (1 - cdf_x)^(n-r)*pdf_x
    prob_dist
}

a = 0
b = 10
r = 5
n = 7

moment_k <- function(k, r, n, a, b) {
    integrand <- function(x) x^k * fr_n(x, r, n, a, b)
    moment <- integrate(integrand, a, b)$value
    moment
}

mean_x5 <- moment_k(1, r, n, a, b)
var_x5 <- moment_k(2, r, n, a, b) - mean_x5^2


cat("Mean of X_{(5)}:", mean_x5, "\n")
cat("Variance of X_{(5)}:", var_x5, "\n")


x_values = seq(a, b, length.out=100)
Fr_n_values = array()
for(i in 1:100) {
  Fr_n_values[i] = integrate(fr_n, a, x_values[i], r=r, n=n, a=a,  b=b)$value 
}

plot(x_values, Fr_n_values, type = "l", xlab = "x->", ylab=paste0("F", r, ":", n, "(x)"))
```

##### Problem 5 A random sample of size 3 is drawn from a population that has a uniform distribution on the interval (0, 1). The resulting order statistics are $X_{1:3}, X_{2:3}$ and $X_{3:3}$. Determine the mean and variance of smallest and largest order statistic.

##### **Method 1:**

```{r}

# kth moment of rth order statistic  = (3!*(k+r-1)!) / ((r-1)! * (k+3)!)
rm_uniform_order <- function(k, r) {
  (factorial(3)*factorial(k+r-1)) / (factorial(r-1)*factorial(k+3))
}

mean_X_1.3 = rm_uniform_order(k = 1, r = 1)
 var_X_1.3 = rm_uniform_order(k = 2, r = 1) - mean_X_1.3^2
mean_X_3.3 = rm_uniform_order(k = 1, r = 3)
 var_X_3.3 = rm_uniform_order(k = 2, r = 3) - mean_X_3.3^2
 
mean_X_1.3
var_X_1.3

mean_X_3.3
var_X_3.3
```

##### **Method 2**

```{r}
# kth moment of rth roder statistic when sample size is n
# PI_i=1^k ((r+i) / (n+i))
rm_uniform_order_n <- function(n, r, k) {
  mult = 1
  for(i in 1:k) {
    term = (r+i-1) / (n+i)
    mult = mult * term
  }
  return(mult)
}
mean_X_1.3 = rm_uniform_order_n(n = 3, r = 1, k = 1)
 var_X_1.3 = rm_uniform_order_n(n = 3, r = 1, k = 2) - mean_X_1.3^2
mean_X_3.3 = rm_uniform_order_n(n = 3, r = 3, k = 1)
 var_X_3.3 = rm_uniform_order_n(n = 3, r = 3, k = 2) - mean_X_3.3^2
 
mean_X_1.3
var_X_1.3

mean_X_3.3
var_X_3.3
```


##### Problem 6: The random sample $X_1, X_2, ..., X_9$ of size 9 is drawn from a population that has a uniform distribution with the interval (0, 10). Evaluate the mean and variance of 7th order statistic.

```{r}
# cdf 
cdf_uniform <- function (x, a, b) {
    ifelse(x < a, 0, ifelse(x > b, 1, (x - a) / (b - a)))
}

# fr:n
fr_n <- function (x, r, n, a, b) {
    cdf_x <- cdf_uniform(x, a, b)
    pdf_x <- 1 / (b - a)

    numerator <- factorial(n) / (factorial(r-1)*factorial(n-r))
    prob_dist <- numerator * cdf_x^(r-1) * (1 - cdf_x)^(n-r)*pdf_x
    prob_dist
}


a <- 0
b <- 10
n <- 9
r <- 7

x_values <- seq(a, b, length.out=100)

fr_n_values <- sapply(x_values, fr_n, r=r, n=n, a=a, b=b)

plot(x_values, fr_n_values, type="l", ylim=c(0, max(fr_n_values)), xlab="x->", ylab=paste0("f", r, ":", n, "(x)"), main=paste0("Probability Density Function of Xr:n"))

# int 2 to 10
p_x7_gt_2 <- integrate(fr_n, 2, b, r=r, n=n, a=a, b=b)$value
cat("P(X_{(7)} > 2):", p_x7_gt_2, "\n")

moment_k <- function(k, r, n, a, b) {
    integrand <- function(x) x^k * fr_n(x, r, n, a, b)
    moment <- integrate(integrand, a, b)$value
    moment
}

mean_x7 <- moment_k(1, r, n, a, b)
var_x7 <- moment_k(2, r, n, a, b) - mean_x7^2


cat("Mean of X_{(7)}:", mean_x7, "\n")
cat("Variance of X_{(7)}:", var_x7, "\n")
```

##### Problem 7: A random sample of size 3 is drawn from a population that has a uniform distribution on the interval (0, 1). The resulting order statistics are $X_{1:3}, X_{2:3}, X_{3:3}$. Determine the mean and variance of sample range. 

```{r}
fr_n = function(x, r, n, a, b) {
  cdf_x = punif(x, min=a, max=b)
  pdf_x = dunif(x, min=a, max=b)
  
  numerator <- factorial(n) / (factorial(r-1)*factorial(n-r))
    prob_dist <- numerator * cdf_x^(r-1) * (1 - cdf_x)^(n-r)*pdf_x
    prob_dist
}

x = seq(0, 1, length.out=100)

a = 0
b = 1
n = 3


moment_k <- function(k, r, n, a, b) {
    integrand <- function(x) x^k * fr_n(x, r, n, a, b)
    moment <- integrate(integrand, a, b)$value
    moment
}

# Expectation of x,y
exp_xy <- function(n, a, b) {
  integrand <- function(x) {
    f_y <- function(y) x * y * n * (n - 1) * (y - x)
    integrate(Vectorize(f_y), lower = x, upper = b)$value
  }
  
  result <- integrate(Vectorize(integrand), lower = a, upper = b)$value
  result
}




mean_x1_n = moment_k(1, r=1, n, a, b)
mean_xn_n = moment_k(1, r=n, n, a, b)

var_x1_n = moment_k(2, r=1, n, a, b) - mean_x1_n^2
var_xn_n = moment_k(2, r=n, n, a, b) - mean_xn_n^2

mean_x1_n
mean_xn_n

var_x1_n
var_xn_n

cov_order = exp_xy(n, a, b) - (mean_x1_n * mean_xn_n)

cov_order

mean_sample_range <- (mean_xn_n - mean_x1_n)
var_sample_range <- (var_x1_n + var_xn_n - 2*cov_order)

cat("Mean of sample range:", mean_sample_range, "\n")
cat("Variance of sample range:", var_sample_range)
```

#### Problem 8: A random sample of size 7 is drawn from a population that has a uniform distribution on the interval (0, 5). The resulting order statistics are $X_{1:7}, X_{2:7}, X_{3:7}, X_{4:7}, X_{5:7}, X_{6:7}$ and $X_{7:7}$. Determine the PDF, mean and variance of $4^{th}$ order (median) statistic. Also find mean and variance of smallest and largest order statistics

```{r}
fr_n = function(x, r, n, a, b) {
  pdf_x <- dunif(x, min=a, max=b)
  cdf_x <- punif(x, min=a, max=b)
  
  numerator <- factorial(n) / (factorial(r-1)*factorial(n-r))
  prob_dis <- numerator * cdf_x^(r-1) * (1 - cdf_x)^(n-r) * pdf_x
  
  return(prob_dis)
}

moment_k <- function(k, r, n, a, b) {
  integrand <- function(x) x^k * fr_n(x, r, n, a, b)
  moment <- integrate(integrand, a, b)$value
  
  return(moment)
}
a = 0
b = 5
n = 7
# median, r = 4
r = n %% 2
x = seq(a, b, length.out = 100)

mean_median <- moment_k(k=1, r, n, a, b)
var_median <- moment_k(k=2, r, n, a, b) - mean_median^2

cat("Mean of median order statistics:", mean_median, "\n")
cat("Variance of median order statistics:", var_median, "\n")

plot(x, fr_n(x, r, n, a, b), type = "l")

mean_x1_n = moment_k(k=1, r=1, n, a, b)
var_x1_n = moment_k(k=2, r=1, n, a, b) - mean_x1_n^2

mean_xn_n = moment_k(k=1, r=n, n, a, b)
var_xn_n = moment_k(k=2, r=n, n, a, b) - mean_xn_n^2

cat("Smallest order statistics: mean =", mean_x1_n, "and variance =", var_x1_n, "\n")
cat("Largest order statistics: mean =", mean_xn_n, "and variance =", var_xn_n, "\n")
```

#### Binomial Distribution

```{r}
set.seed(11)
par(mfrow=c(1, 3))
n = c(9, 9, 9) # size
p = c(0.55, 0.35, 0.4) # Probability
color = c("red", "green", "blue")
for(i in 1:3) {
 x = rbinom(1000, n[i], p[i])
 hist(x, col=color[i], main = paste("Histogram with n =", n[i], "and p=", p[i]))
}
```


#### Poisson Distribution

```{r}
set.seed(12)
par(mfrow=c(1,3))
p <- c(3,5,9)
color <- c("red", "green", "blue")
for(i in 1:3) {
 x = rpois(1000, p[i])
 hist(x, col=color[i], main=paste("Poisson with lemda =", p[i]))
}
```